TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-13 paper/draft.tex
keywords

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution Apache Point Galactic Evolution Experiment (\apogee; near-infrared) spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H]) to precisions higher than the model-grid pipeline.
The lack of reliable stellar parameters reported by the \apogee\ pipeline for temperatures less than $\sim$3500K motivates the extension of this approach to M dwarf stars, but with new training labels.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 68\,K and 0.07\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging from M0 to M9, and demonstrate that \thecannon\ acting on \apogee\ spectra can infer spectral types with predictive accuracy of 0.6 types.
The trained models are employed to determine spectral subtypes, temperatures and metallicities for 10,311 M dwarfs in \apogee.
We validate our results using photometric temperature indicators and the \gaia\ color--magnitude diagram.
As a side result, we determine empirical relationships between spectral type and temperature and metallicity.
These results have implications for the \tess\ Mission and other projects that involve large samples of M dwarfs.
\end{abstract}

\keywords{
surveys
 ---
stars: Mdwarfs
 ---
stellar parameters
 ---
methods:~data~analysis
}

>>>>

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution Apache Point Galactic Evolution Experiment (\apogee; near-infrared) spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H]) to precisions higher than the model-grid pipeline.
The lack of reliable stellar parameters reported by the \apogee\ pipeline for temperatures less than $\sim$3500K motivates the extension of this approach to M dwarf stars, but with new training labels.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 68\,K and 0.07\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging from M0 to M9, and demonstrate that \thecannon\ acting on \apogee\ spectra can infer spectral types with predictive accuracy of 0.6 types.
The trained models are employed to determine spectral subtypes, temperatures and metallicities for 10,311 M dwarfs in \apogee, selected in part on \gaia\ photometry and astrometry.
We validate our results using photometric temperature indicators and the \gaia\ color--magnitude diagram.
As a side result, we determine empirical relationships between spectral type and temperature and metallicity.
These results have implications for the \tess\ Mission and other projects that involve large samples of M dwarfs.
\end{abstract}

\keywords{
infrared:~stars
 ---
methods:~data~analysis
 ---
stars:~abundances
 ---
stars:~fundamental~parameters
 ---
stars:~late-type
 ---
surveys
}
